Project 2
==========

## 1) load ToothGrowth dataset and perform exploratory analysis
```{r}
library(ggplot2)
library(plyr)

tooth <- ToothGrowth
tooth.agg2 <- aggregate(tooth$len, list(supp = tooth$supp, dose = tooth$dose), mean)
names(tooth.agg2)[3] <- "len"
ggplot(tooth.agg2, aes(x = dose, y = len, fill = supp)) + geom_bar(stat = "identity", position = position_dodge()) + 
        ggtitle("Comparison of average len for OJ and VC doses")

tooth.freq <- table(tooth$supp, tooth$dose)
tooth.freq <- as.data.frame(tooth.freq)
names(tooth.freq) <- c("supp", "dose", "frequency")
ggplot(tooth.freq, aes(x = dose, y = frequency, fill = supp)) + geom_bar(stat = "identity", position = position_dodge()) + ggtitle("Frequency of dosage by supp")
```        
## 2) provide a basic summary of the data

## summary function called on data
```{r}
summary(tooth)
```

## mean len for OJ and VC
```{r}
aggregate(tooth$len, list(supp = tooth$supp), mean)
```

## descriptive statistics for VC and OJ by dosage
```{r}
ddply(tooth, c("supp", "dose"), summarise, n = length(len), mean = mean(len), sd = sd(len), se = sd/sqrt(n))
```

## 3) use confidence intervals and hypothesis tests to compare tooth growth by supp and dose

## compare mean length of VJ and OJ
```{r}
t.test(len ~ supp, data = tooth)
```
## since p-value is 0.06, we cannot reject the null hypothesis

## compare mean length of dosages .5 and 1
```{r}
tooth.subset <- subset(tooth, tooth$dose != 2.0)
t.test(len ~ dose, data = tooth.subset)
```
## can reject null hypothesis

## compare mean length of dosages .5 and 2
```{r}
tooth.subset <- subset(tooth, tooth$dose != 1.0)
t.test(len ~ dose, data = tooth.subset)
```
## can reject null hypothesis

## compare mean length of dosages 1 and 2
```{r}
tooth.subset <- subset(tooth, tooth$dose != 0.5)
t.test(len ~ dose, data = tooth.subset)
```
## can reject null hypothesis

## compare mean length of dosage .5 for VC and OJ
```{r}
tooth.subset <- subset(tooth, tooth$dose == 0.5)
t.test(len ~ supp, data = tooth.subset)
```
## can reject null hypothesis

## compare mean length of dosage 1 for VC and OJ
```{r}
tooth.subset <- subset(tooth, tooth$dose == 1.0)
t.test(len ~ supp, data = tooth.subset)
```
## can reject null hypothesis

## compare mean length of dosage 2 for VC and OJ
```{r}
tooth.subset <- subset(tooth, tooth$dose == 2.0)
t.test(len ~ supp, data = tooth.subset)
```
## cannot reject null hypothesis

## 4) state conclusions and assumptions

## the main finding is:
## for dosages of .5 and 1, the mean of OJ is higher than the mean of VC, 
## and the difference in means is both statistically and practically significant

## the main assumption is that the observations in the ToothGrowth dataset are independent, and not paired
